---
title: "BIOL432_A6"
author: "Jennah Levac"
date: "2025-02-10"
output: html_document
---

```{r}
library(doParallel)
library(parallel)
library(foreach)
```
Laptop with 8 Cores
```{r}
detectCores()
```
```{r}
StartTime <- Sys.time()

for (i in 1:4000) {
  random_numbers <- rnorm(100000, mean = 10, sd = 3)
  mean_random_numbers <- mean(random_numbers)}

EndTime <- Sys.time()

TotalTime <- EndTime - StartTime

print(TotalTime)
```
Therefore, initial run time was estimated be ~25.7sec
```{r}
E_V<- 1
NCores<-detectCores()
```

```{r}
Estimate<- TotalTime/(E_V * (NCores-1))
print(Estimate)
```
```{r}
Cores <- parallel::makeCluster(detectCores() - 1)  
```
```{r}
doParallel::registerDoParallel(Cores)
```
```{r}
Task1 <- function(n) {
  random_numbers2 <- rnorm(n, mean = 10, sd = 3)
  return(mean(random_numbers2))}
```
```{r}
StartTime2 <- Sys.time()
Results<- foreach(i = 1:4000, .combine = rbind) %dopar% {
  Task1(100000)}

EndTime2 <- Sys.time()

TotalTime2 <- EndTime2 - StartTime2
```
```{r}
print(TotalTime2)
```
```{r}
parallel::stopCluster(Cores)
```
Due to internal processes, the actual parallel loop time is higher (~7.6sec) than our theoretical time (~3.6), which is calculated "under perfect conditions". However, there is still significant time improvement in the parallel loop versus serial loop (~25.7sec), since the parallel loop allows tasks to be divided and completed at the same time (i.e. parallel to one another) and therefore returning results quicker.